### 2013-09-09 thesis/src/chapters/chapter-4.tex
Add clients' view on cloud computing section - continued

\subsection{Current usage and trends}
\subsubsection*{Cloud -- clients' view}
Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing. As the recent survey \cite{RackspaceSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general. However, the tendency is that more and more enterprises do not find it a major issue as in 2012 the number declined to 46\%.
Complexity related to the managament of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obsctacles to adoption in 2013 for they constitued 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively. Total results are shown in the figure \ref{ch4:cloud-computing-adoption-obstacles-2013}.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-4/cloud-computing-adoption-obstacles-2013}
  \end{center}
  \caption{Major obstacles to cloud adoption in 2013 according to \cite{RackspaceSurvey13}}
  \label{ch4:cloud-computing-adoption-obstacles-2013}
\end{figure}

>>>>

\subsection{Current usage and trends}
\subsubsection*{Cloud -- clients' view}
Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing. As the recent survey \cite{NBSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general. However, the tendency is that more and more enterprises do not find it a major issue as in 2012 the number declined to 46\%.
Complexity related to the management of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obstacles to adoption in 2013 for they constituted 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively. Total results are shown in the figure \ref{ch4:cloud-computing-adoption-obstacles-2013}.
\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-4/cloud-computing-adoption-obstacles-2013}
  \end{center}
  \caption{Major obstacles to cloud adoption in 2013 according to \cite{NBSurvey13}}
  \label{ch4:cloud-computing-adoption-obstacles-2013}
\end{figure}

The same survey shows that the cloud adoption growth rate is high -- 75 percent of responders stated usage of some sort of a cloud platform. This means 8 percent growth when compared to the results obtained in 2012. The expectations for the total worldwide addressable market for cloud computing are to reach \$158.8B by 2014 -- an increase of 126.5 percent from 2011.
